Effort and Delay Discounting in a Foraging Environment

### 2019-05-08 Cost2/Cost2_prereg_remake.Rmd
done with mix logistic, but revisit the theory

``` {r 16.2.3. Mixed Effects Model, echo=FALSE,fig.align="center",fig.width=5,fig.height=4}
mixLogis_main <- list()
mixData <- dataAll %>%
  filter(Choice < 2) %>%
  group_by(SubjID, Cost, Handling, Reward) %>%
  summarize(totalChoices = length(Choice),
            totalAccepted = sum(Choice),
            totalQuits = sum(Choice == 0),
            propAccepted = mean(Choice)) %>%
  ungroup()

testBinom <- cbind(mixData$totalAccepted, mixData$totalQuits)

mixData <- within(mixData, Cost <- relevel(Cost, ref = "Wait"))
mixLogis_main$Wait <-  glmer(cbind(totalAccepted, totalQuits) ~ Cost + Handling + Reward + (1 | SubjID), family = "binomial", data = mixData)
mixData <- within(mixData, Cost <- relevel(Cost, ref = "Cognitive"))
mixLogis_main$Cognitive <-  glmer(cbind(totalAccepted, totalQuits) ~ Cost + Handling + Reward + (1 | SubjID), family = "binomial", data = mixData)
mixData <- within(mixData, Cost <- relevel(Cost, ref = "Physical"))
mixLogis_main$Physical <-  glmer(cbind(totalAccepted, totalQuits) ~ Cost + Handling + Reward + (1 | SubjID), family = "binomial", data = mixData)
mixData <- within(mixData, Cost <- relevel(Cost, ref = "Easy"))
mixLogis_main$Easy <-  glmer(cbind(totalAccepted, totalQuits) ~ Cost + Handling + Reward + (1 | SubjID), family = "binomial", data = mixData)

temp <- lapply(mixLogis_main, function(model) {coefficients(model)$SubjID[1, 2:4]})
mixCoeffs <- bind_rows(temp) #reduce(temp, full_join) 
dimnames(mixCoeffs) <- list(names(temp), substr(colnames(mixCoeffs), 5, 20))
mixCoeffs <- as.matrix(mixCoeffs[, match(rownames(mixCoeffs), colnames(mixCoeffs))])
mixCoeffs[is.na(mixCoeffs)] <- 1
mixCoeffs <- mixCoeffs[c("Cognitive", "Physical", "Wait", "Easy"), c("Cognitive", "Physical", "Wait", "Easy")]

# # just to make the logistic plotting comply with the function
# allProp2 <- allProp2[order(as.character(allProp2$Group), decreasing = F), ]
# 
# # This transformation is necessary for a proper binomial fit of proportion data
# totalQuit <- allProp2$totalTrials - allProp2$totalComplete
# testBinom <- cbind(allProp2$totalComplete,totalQuit)
# 
# # Using a mixed-model fitted with GLMER
# # Note: GLMER does not take quasi family distributions
# # ranef() to display the random intercepts, fixef() for the model coefficients
# 
# # fit model
# mixLogis_main <- list()
# allProp2 <- within(allProp2, Group <- relevel(Group, ref = "Wait"))
# mixLogis_main$Wait <-  glmer(testBinom ~ Group + Handling + Reward + (1 | SubjID), family = "binomial", data = allProp2)
# allProp2 <- within(allProp2, Group <- relevel(Group, ref = "Cognitive"))
# mixLogis_main$Cognitive <-  glmer(testBinom ~ Group + Handling + Reward + (1 | SubjID), family = "binomial", data = allProp2)
# allProp2 <- within(allProp2, Group <- relevel(Group, ref = "Physical"))
# mixLogis_main$Physical <-  glmer(testBinom ~ Group + Handling + Reward + (1 | SubjID), family = "binomial", data = allProp2)
# allProp2 <- within(allProp2, Group <- relevel(Group, ref = "Easy"))
# mixLogis_main$Easy <-  glmer(testBinom ~ Group + Handling + Reward + (1 | SubjID), family = "binomial", data = allProp2)
# 
# # Summarize model 
# mixLogis_main$Summary <- summary(mixLogis_main$Wait)
# 
# # Coefficient matrices for all reference combos
# t1 <- c(summary(mixLogis_main$Wait)$coefficients[2:4,1], GroupWait = 0)
# t2 <- c(summary(mixLogis_main$Cognitive)$coefficients[2:4,1], GroupCognitive = 0)
# t3 <- c(summary(mixLogis_main$Physical)$coefficients[2:4,1], GroupPhysical = 0)
# t4 <- c(summary(mixLogis_main$Easy)$coefficients[2:4,1], GroupEasy = 0)
# mixLogis_mainBetas <- cbind(merge(t1,t2, by = 0), merge(t1,t3,by=0)[,3], merge(t1,t4,by=0)[,3])
# dimnames(mixLogis_mainBetas) <- list(cond[c(2,4,3,1)],c("Group",cond))
# mixLogis_mainBetas <- as.matrix(mixLogis_mainBetas[,c(3,5,4,2)])
# 
# # And their respective p-values
# t1 <- c(summary(mixLogis_main$Wait)$coefficients[2:4,4], GroupWait = 1)
# t2 <- c(summary(mixLogis_main$Cognitive)$coefficients[2:4,4], GroupCognitive = 1)
# t3 <- c(summary(mixLogis_main$Physical)$coefficients[2:4,4], GroupPhysical = 1)
# t4 <- c(summary(mixLogis_main$Easy)$coefficients[2:4,4], GroupEasy = 1)
# mixLogis_mainPvals <- cbind(merge(t1,t2, by = 0), merge(t1,t3,by=0)[,3], merge(t1,t4,by=0)[,3])
# dimnames(mixLogis_mainPvals) <- list(cond[c(2,4,3,1)],c("Group",cond))
# mixLogis_mainPvals <- as.matrix(mixLogis_mainPvals[,c(3,5,4,2)])
# 
# # Create CIs
# sd <- sqrt(diag(vcov(mixLogis_main$Wait)))
# mixLogisCI <- cbind(Est = fixef(mixLogis_main$Wait), LL = fixef(mixLogis_main$Wait) - 1.96 * sd, UL = fixef(mixLogis_main$Wait) + 1.96 * sd)
```

@@

# plot per subject (as expected, the intercepts are highest for those who accepted everythiing)
temp <- cbind(summaryData$Group, 
              ranef(mixLogis_main$Wait)$SubjID)#, 
              #Colors = c(rep(cols[1],nSubsW),rep(cols[2],nSubsC),rep(cols[3],nSubsP),rep(cols[4],nSubsE)))
colnames(temp) <- list("Group", "RandIntercept")
RI1 <- ggplot(data = temp, aes(x = Group, y = RandIntercept, fill = Group)) + 
  geom_hline(yintercept = 0, alpha = 0.7) + 
  geom_boxplot(show.legend = F) +
  scale_fill_manual(values = cols) + 
  labs(title = "Random Intercepts per Group", x = "") +
  theme_classic()

RI1
```

>>>>

``` {r 16.2.3. Mixed Effects Model, echo=FALSE,fig.align="center",fig.width=5,fig.height=4}
# clean the data so the total acceptances and quits per subj x handling x reward are ready for modeling
mixLogis_main <- list()
mixData <- dataAll %>%
  filter(Choice < 2) %>%
  group_by(SubjID, Cost, Handling, Reward) %>%
  summarize(totalChoices = length(Choice),
            totalAccepted = sum(Choice),
            totalQuits = sum(Choice == 0),
            propAccepted = mean(Choice)) %>%
  ungroup()

# model with a random intercept and cost-type slope, then relevel the cost to get all pairwise comparisons
# the correct way to model a logistic with proportions as dependent vars is by prividing n of hits and quits as a matrix
mixData <- within(mixData, Cost <- relevel(Cost, ref = "Wait"))
mixLogis_main$Wait <-  glmer(cbind(totalAccepted, totalQuits) ~ Cost + Handling + Reward + (Cost | SubjID), family = "binomial", data = mixData)
mixData <- within(mixData, Cost <- relevel(Cost, ref = "Cognitive"))
mixLogis_main$Cognitive <-  glmer(cbind(totalAccepted, totalQuits) ~ Cost + Handling + Reward + (Cost | SubjID), family = "binomial", data = mixData)
mixData <- within(mixData, Cost <- relevel(Cost, ref = "Physical"))
mixLogis_main$Physical <-  glmer(cbind(totalAccepted, totalQuits) ~ Cost + Handling + Reward + (Cost | SubjID), family = "binomial", data = mixData)
mixData <- within(mixData, Cost <- relevel(Cost, ref = "Easy"))
mixLogis_main$Easy <-  glmer(cbind(totalAccepted, totalQuits) ~ Cost + Handling + Reward + (Cost | SubjID), family = "binomial", data = mixData)

# matrix of coefficients
temp <- lapply(mixLogis_main, function(model) {coefficients(model)$SubjID[1, 2:4]})
mixCoeffs <- bind_rows(temp) #reduce(temp, full_join) 
dimnames(mixCoeffs) <- list(names(temp), substr(colnames(mixCoeffs), 5, 20))
mixCoeffs <- as.matrix(mixCoeffs[, match(rownames(mixCoeffs), colnames(mixCoeffs))])
mixCoeffs[is.na(mixCoeffs)] <- 1
mixCoeffs <- mixCoeffs[c("Cognitive", "Physical", "Wait", "Easy"), c("Cognitive", "Physical", "Wait", "Easy")]

# matrix of pvals
temp <- lapply(mixLogis_main, function(model) {as.list(summary(model)$coefficients[2:4, 4])})
mixPvals <- as.matrix(bind_rows(temp)) 
dimnames(mixPvals) <- list(names(temp), substr(colnames(mixPvals), 5, 20))
mixPvals <- as.matrix(mixPvals[, match(rownames(mixPvals), colnames(mixPvals))])
rownames(mixPvals) <- colnames(mixPvals)
mixPvals <- mixPvals[c("Cognitive", "Physical", "Wait", "Easy"), c("Cognitive", "Physical", "Wait", "Easy")]

# plot coefficients and pvalues in a matrix
# invert colors
col2 <- colorRampPalette(c("#053061", "#2166AC", "#4393C3", "#92C5DE", "#D1E5F0", "#FFFFFF", "#FDDBC7", "#F4A582", "#D6604D", "#B2182B", "#67001F"))

corrplot(mixCoeffs, 
         is.corr = F, 
         p.mat = mixPvals, 
         type = "upper",
         outline = T,
         insig = "p-value",
         sig.level = -1, 
         na.label = "square", 
         na.label.col = "grey",
         method = "color", 
         tl.col = "black",
         tl.cex = 0.8,
         col = col2(200))

@@

# plot per subject (as expected, the intercepts are highest for those who accepted everythiing)
temp <- ranef(mixLogis_main$Wait)$SubjID %>%
  gather(key = Cost, value = RandEffects, colnames(.))

(RI1 <- ggplot(data = temp, aes(Cost, RandEffects, fill = Cost)) + 
  geom_hline(yintercept = 0, alpha = 0.7) + 
  geom_boxplot(show.legend = F) +
  scale_fill_manual(values = cols) + 
  labs(x = "") +
  theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank(), 
              panel.background = element_blank(), 
              axis.line = element_line(colour = "black"),
              text = element_text(size = 16)))
